Human judgement forecasting of COVID-19 in the UK

### 2023-05-04 manuscript/manuscript.tex
Some manuscript corrections

\title{Human Judgement Forecasting of COVID-19 in the UK}
%\titlenote{The title should be detailed enough for someone to know whether the article would be of interest to them, but also concise. Please ensure the broadness and claims within the title are appropriate to the content of the article itself.}

@@

From May 24th to August 16th 2021, we elicited weekly one to four week ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. A median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$. This was then used to forecast cases and deaths using simulation methods from the \texttt{EpiNow2} \textsf{R} package. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

@@

Human judgement forecasts and computational models can produce forecasts of similar quality for infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a "good" forecast is dependent on the forecast consumer. Combinations of human and computational forecasts hold potential but present real-world challenges that need to be solved.

@@

Past research found that in the context of infectious disease forecasting, human judgement forecasts could achieve predictive performance broadly comparable to forecasts generated based on mathematical modelling, in particular when forecasting incident cases, rather than lagged indicators indicators like deaths. \citet{farrowHumanJudgmentApproach2017} found that an aggregate of human predictions outperformed computational models when predicting the 2014/15 and 2015/16 flu season in the US. However, a comparable approach performed worse than computational models at predicting the 2014/15 outbreak of chikungunya in the Americas. 
\citet{bosseComparingHumanModelbased2022} found an ensemble of human forecasters to outperform an ensemble of computational models when predicting cases of COVID-19 in Germany and Poland, but performing worse when predicting incident deaths. Similarly, \citet{mcandrewChimericForecastingCombining2022} reported an ensemble of human forecasters to perform comparably to an ensemble of computational models when predicting incident COVID-19 cases, and worse when predicting incident deaths. \citet{farrowHumanJudgmentApproach2017} and in particular \citet{bosseComparingHumanModelbased2022} struggled to recruit a large number of participants (numbers of active forecasters ranged from 22 to 61 in \citet{mcandrewChimericForecastingCombining2022}, 7 to 24 in \citet{farrowHumanJudgmentApproach2017}, and 4 to 10 in \citet{bosseComparingHumanModelbased2022}). 
It is important to note that in previous studies (and also this one) human forecasters were free to use any resources, including computational models, in the process of creating a forecast, making it difficult to completely separate human judgement and computational modelling.

@@

This paper represents a follow-up study to \citet{bosseComparingHumanModelbased2022} in the United Kingdom with forecasts made over the course of thirteen weeks between May and August 2021. Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2022}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country, in a different time period, and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the estimated effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a simulation model. We describe this approach as human in the loop computational modelling and consider it a formalisation of often practiced manual intervention in computational forecasts.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022a} was launched in March 2021 in order to elicit weekly predictions for various COVID-19 related forecast targets from different research groups. The forecasts evaluated in this study were submitted every Monday before 11.59pm GMT between May 24 2021 and August 16 2021. Forecasts were made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epidemiological weeks, ending on a Saturday and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub followed a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecasts submitted to the hub were automatically checked for conformity with the required format and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble ("EuroCOVIDhub-ensemble") shown on all official Forecast Hub visualisations (https://covid19forecasthub.eu/) was a mean-ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July onwards, the default Forecast Hub ensemble became a median ensemble. The median number of models included in the Forecast Hub ensemble for the UK during the study period was 9 for cases and 10 for deaths (see Figure \ref{fig:num-forecasters} in the SI).

@@

Just as in the previous study, the default forecast shown was a repetition of the last known observation with constant uncertainty around it. The shown distribution was the exponential of a normal distribution with mean $\log (\text{last value})$ and uncertainty set to the standard deviation of the last four changes in weekly log observed forecasts (i.e. as $\sigma(\log(\text{value4}) - \log(\text{value3}), \, \log(\text{value3}) - \log(\text{value2}), \, \dots)$).
In addition to information about past observations, participants could see various metrics and data such as the test positivity rate and vaccination rate sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

In addition to the 'direct' forecasts, we implemented a second forecasting method ('$R_t$ forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$. This forecast was made based on a baseline estimate produced by the \texttt{EpiNow2} \citep{epinow2} \textsf{R} \cite{R} package effective reproduction number model which we also used in \cite{bosseComparingHumanModelbased2022} as a standalone computational model. The estimate produced by \texttt{EpiNow2} was shown as the default forecast and could be adjusted by the user. The resulting $R_t$ forecast was then translated into a forecast of cases using the simulation model from the \texttt{EpiNow2} \textsf{R} package, which implements a renewal equation based \citep{fraserEstimatingIndividualHousehold2007} generative process for latent infections. We chose a Gaussian Process prior with mean 0 for the first differences of the effective reproduction number in time, implying that in the absence of informative data the reproduction number would remain constant on average, with uncertainty increasing with the temporal distance to informative data points. Latent infections were convolved with delay distributions representing the incubation period and reporting delay, and assumed to follow a negative binomial observation model with a day of the week effect to produce an estimate of reported cases. This approach has been widely used for short-term forecasting \cite{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a} and used to produce reproduction number estimates \cite{abbottEstimatingIncreaseReproduction2021, abbottEstimatingTimevaryingReproduction2020a, sherrattExploringSurveillanceData2021}. Further details are given in the SI.

In order to obtain forecasts for deaths, we similarly fit a model that convolved observed and predicted reported cases as implied by the $R_t$ forecast over a  delay distributions \cite{sherrattExploringSurveillanceData2021, abbottEstimatingTimevaryingReproduction2020a} and scaled them by a fixed ratio to model the time between a case report and a reported death and the case fatality ratio using the \texttt{EpiNow2} \textsf{R} package \citep{epinow2}. Further details are given in the SI and in our previous work \citep{bosseComparingHumanModelbased2022}.

@@

In order to boost participation compared to our last crowd forecasting study in Germany and Poland \citep{bosseComparingHumanModelbased2022} which struggled in this regard, we announced an official tournament, the "UK Crowd Forecasting Challenge". Participants were asked to submit weekly predictions for reported cases and deaths linked to COVID-19 in the United Kingdom one to four weeks into the future. Everyone who had submitted a forecast for targets in the UK during the tournament period from the 24th of May 2021 to the 16th of August 2021 was deemed a participant and eligible for a prize. The first prize was 100 GBP, second prize 50 GBP and third prize 25 GBP. Participant performance was determined using the mean weighted interval score (WIS) on the log scale (see details in the next Section), %\ref{sec:analysis}), 
averaged across forecast dates, horizons and forecast targets. For the purpose of the tournament ranking, participants who did not submit a forecast in a given week were assigned the median score of all other participants who submitted a forecast that week. The UK crowd forecasting challenge was announced over Twitter and our networks. 
% (e.g. https://www.voice-global.org/public/opportunities/archived/uk-covid-19-crowd-forecasting-challenge/)
In addition, we created a project website, \url{https://crowdforecastr.org}, made weekly posts on Twitter and sent participants who had registered on the online application weekly emails with a reminder and a summary of their past performance. A public leaderboard was available on our website \url{https://epiforecasts.io}. Participants could choose to make a direct forecast as well as an $R_t$ forecast and were counted as two separate forecasters and eligible for prizes twice. Weekly forecasts had to be submitted between Sunday 12pm and Monday 8pm UK time.

@@

Our analysis of the data largely follows the one outlined in \citet{bosseComparingHumanModelbased2022}. We scored forecasts using the weighted interval score (WIS, formulas given in the SI) 
\cite{bracherEvaluatingEpidemicForecasts2021}. The WIS is a strictly proper scoring rule yielding non-negative values, with smaller values implying better performance. A forecaster, in expectation, optimises their score by providing a predictive distribution $F$ that is equal to the data-generating distribution $G$, and is therefore incentivised to report their true belief. The WIS can be understood as an approximation of the continuous ranked probability score (CRPS) for forecasts in a quantile-based format, which itself represents a generalisation of the absolute error to predictive distributions. The WIS can be decomposed into three separate penalty components: forecast dispersion (i.e. uncertainty of forecasts), overprediction and underprediction.

\citet{bosseTransformationForecastsEvaluating2023} recently suggested to transform forecasts and observations using the natural logarithm prior to applying the WIS in order to better reflect the exponential nature of the underlying disease process. We, therefore, also compute WIS values after transforming all forecasts and observations using the function $f\colon x \to \log (x + 1)$. In the following, we refer to WIS scores obtained without a transformation as "scores on the natural scale", and WIS values obtained after log-transforming forecasts and observations as "scores on the log scale". To make scores easier to interpret, we report relative WIS scores, where the average score for a given model was divided by the average score for the European Forecast Hub ensemble ("EuroCOVIDhub-ensemble"). In addition, we computed ranks based on WIS values.

In order to measure probabilistic calibration \citep{gneitingProbabilisticForecastsCalibration2007}, we used the empirical coverage of all central 50\% and 90\% prediction intervals. Empirical coverage refers to the percentage of observations falling inside any given central prediction interval (e.g. the cumulative percentage of observed values that fall inside all central 50\% prediction intervals).

If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs, which found predictive performance to be poor and unreliable beyond this horizon \cite{cramerEvaluationIndividualEnsemble2021, sherrattPredictivePerformanceMultimodel2022a, bracherShorttermForecastingCOVID192021}. We analysed all forecasts stratified by forecast target (cases or deaths), forecast horizon, and forecast approach. We compared the performance of the direct vs. $R_t$ forecasting approach using instances where we had both a direct forecasts and an $R_t$ forecast from the same person.

@@

A total number of 90 participants submitted forecasts (more precisely, forecasts were submitted from 90 different accounts, some of them anonymous). Out of 90 participants, 21 self-identified as "experts", i.e. stated they had professional experience in infectious disease modelling or a related field.

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/scores-and-forecasts.png}
\caption{\bf{Forecasts and corresponding WIS for 2-week ahead forecasts of cases and deaths from COVID-19 in the UK.} A: 50\% prediction intervals (coloured bars) and observed values (black line and points) for cases and deaths on the natural scale. B: Corresponding WIS values, decomposed into dispersion, overprediction and underprediction. C: 50\% prediction intervals on the log scale, i.e. after applying the natural logarithm to all forecasts and observations. D: Corresponding WIS on the log scale, i.e. the WIS applied to the log-transformed forecasts and observations.} 
\label{fig:forecasts-scores} 
\end{figure*}

@@

\subsection*{Experts and Non-Experts}

@@

%Calibration
In line with results from previous work \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022a}, we found almost all forecasts for cases to be underdispersed (i.e. too narrow/overconfident). Empirical coverage for death forecasts was higher than the corresponding coverage for cases for all forecasting approaches except the $R_t$ forecasts.

@@

This study explored a novel method of forecasting infectious diseases that combines a human forecast of the estimated effective reproduction number $R_t$ with epidemiological modelling to map the $R_t$ forecast to a forecast of cases and deaths. One appeal of this approach is that the forecaster can directly forecast the generative process and how they believe it is affected by interventions and changes in behaviour. Computational modelling then takes care of dealing with details such as reporting delays, generation intervals, day of the week periodicity, and the relationship between different indicators. This could help reduce cognitive load, and make it easier to synthesise various sources in information into a single forecast, at least for forecasters who have an intuitive understanding of $R_t$. Though we note all of these modelling steps and the construction of the model itself requires the human constructing the model to make assumptions. Anecdotally, forecasters familiar to the authors reported high satisfaction with the forecasting experience. In our study, $R_t$ forecasts of cases were comparable to direct forecasts, with a tendency for good forecasts to improve when using the $R_t$ method and worse forecasts to deteriorate even more. Sample sizes, however, were very low. 
Given that forecasters could simulate cases in the app, it is also possible that forecasters were in reality directly forecasting cases. 
$R_t$ forecasts of deaths (which forecasters could not see in the app) were noticeably worse than direct forecasts of deaths. The computational model underlying our $R_t$ forecasts of deaths estimated a constant CFR and delay distribution using the last 4 weeks of data, therefore updating relatively slowly to new circumstances and was assumed to be constant over the four week forecast horizon. However, as mentioned before, the CFR evolved during the study period. Forecasters had no way of inspecting the death forecast implied by their $R_t$ forecast and also had no way to adjust the CFR manually, likely impacting forecast accuracy. Giving human forecasters the ability to adjust the CFR and other model parameters would have increased complexity of the interface, but would have solved issues with the assumptions of the underlying model. Alternatively, a more complex model could have been used which allowed for time-varying CFR estimates and forecast these changes over the forecast horizon though this approach may still have struggled to cope with the rapid changes observed during the study period. Another important limitation is that we didn't have full sample trajectories of the $R_t$-values predicted by forecasters. Rather, trajectories had to be constructed based on the distributions provided for the different forecast horizons, which likely negatively affected forecasts. One potential way to disentangle the effect of the convolution model from the $R_t$ forecasts would have been to use the human forecasts for cases as an input to the second computational model, which could then have simulated deaths. Future work could expose forecasters to different combinations of these options with the aim of separating effects of the user interface from ones related to the structure of the underlying computational model.

@@

Past expert forecasts of COVID-19 \cite{recchiaHowWellDid2021} had found predictions from experts to outperform those of non-experts. In our study, an ensemble of self-reported "experts" outperformed an ensemble of "non-experts" when forecasting cases two weeks ahead, both on the log scale and on the natural scale. When forecasting deaths two weeks ahead, "experts" performed worse than "non-experts" on the log scale, but better on the natural scale. Forecasts for "experts" tended to be better calibrated than non-experts.  However results should be taken with care in light of relatively low sample sizes (median of 11 "non-experts" for cases and 10 for deaths, median of 8 "experts" for cases and deaths), and given that expert status was self-reported. Furthermore, we only asked for professional involvement in a field related to infectious disease modelling, not specifically for familiarity with modelling of COVID-19 in the UK, and only offered participants a binary choice. However, as we used ad-hoc recruitment in our networks many of these self-identified experts are likely to be infectious disease modellers.

% \paragraph{Susceptibility of results}
Overall, results of our study should be taken with caution due to a number of important limitations. Firstly, our study was restricted to one location and to a relatively short period of thirteen weeks. Secondly, there were many confounding factors that likely influence results. These include the fact that different participants made forecasts at different points in time (with the median forecaster only submitting a single forecast) and that subgroups of interest (e.g. "experts", or $R_t$ forecasts) had different numbers of forecasters. 
In most instances, differences in scores between forecast approaches were small compared to the variance of scores within a single approach. In addition, there were many researcher degrees of freedom that could influence findings, for example how individual forecasts were combined to create an ensemble. Results were influenced by choices made during the evaluation with, for example, some conclusions  depending on forecast horizon and the transformation used prior to scoring. 
Highlighting this, prizes to the human forecasters were paid out based on the combined WIS on the log scale across all horizons and forecast targets. Had we chosen to instead measure WIS on the natural scale, or to forecast only cases and continue to score on the log scale, rankings and payouts would have been different.

@@

\subsection*{Author contributions}
In order to give appropriate credit to each author of an article, the individual
contributions of each author to the manuscript should be detailed in this section. We
recommend using author initials and then stating briefly how they contributed.

\subsection*{Competing interests}
Nikos I. Bosse is a part-time employee of Metaculus, an online prediction platform.

\subsection*{Grant information}
Please state who funded the work discussed in this article, whether it is your employer,
a grant funder etc. Please do not list funding that you have that is not relevant to this
specific piece of research. For each funder, please state the funder’s name, the grant
number where applicable, and the individual to whom the grant was assigned.

\subsection*{Acknowledgements}
We thank all forecasters for their participation and want to congratulate the three winners of the forecasting challenge: Russel Bradshaw, Sebastian Funk (an author of this study), and Akira Endo. All winners donated their prizes.

>>>>

\title{Human judgement forecasting of COVID-19 in the UK}
%\titlenote{The title should be detailed enough for someone to know whether the article would be of interest to them, but also concise. Please ensure the broadness and claims within the title are appropriate to the content of the article itself.}

@@

From May 24th to August 16th, 2021, we elicited weekly one to four week ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. A median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$. This was then used to forecast cases and deaths using simulation methods from the \texttt{EpiNow2} \textsf{R} package. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

@@

Human judgement forecasts and computational models can produce forecasts of similar quality for infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a ``good" forecast is dependent on the forecast consumer. Combinations of human and computational forecasts hold potential but present real-world challenges that need to be solved.

@@

Past research found that in the context of infectious disease forecasting, human judgement forecasts could achieve predictive performance broadly comparable to forecasts generated based on mathematical modelling, in particular when forecasting incident cases, rather than lagged indicators indicators like deaths. \citet{farrowHumanJudgmentApproach2017} found that an aggregate of human predictions outperformed computational models when predicting the 2014/15 and 2015/16 flu season in the US. However, a comparable approach performed worse than computational models at predicting the 2014/15 outbreak of chikungunya in the Americas. 
\citet{bosseComparingHumanModelbased2022} found an ensemble of human forecasters to outperform an ensemble of computational models when predicting cases of COVID-19 in Germany and Poland, but performing worse when predicting incident deaths. Similarly, \citet{mcandrewChimericForecastingCombining2022} reported an ensemble of human forecasters to perform comparably to an ensemble of computational models when predicting incident COVID-19 cases, and worse when predicting incident deaths. \citet{farrowHumanJudgmentApproach2017} and in particular \citet{bosseComparingHumanModelbased2022} struggled to recruit many participants (numbers of active forecasters ranged from 22 to 61 in \citet{mcandrewChimericForecastingCombining2022}, 7 to 24 in \citet{farrowHumanJudgmentApproach2017}, and 4 to 10 in \citet{bosseComparingHumanModelbased2022}). 
It is important to note that in previous studies (and also this one) human forecasters were free to use any resources, including computational models, in the process of creating a forecast, making it difficult to completely separate human judgement and computational modelling.

@@

This paper represents a follow-up study to \citet{bosseComparingHumanModelbased2022} in the United Kingdom with forecasts made over the course of thirteen weeks between May and August 2021. Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2022}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country, in a different time period, and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the estimated effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a simulation model. We describe this approach as human in the loop computational modelling and consider it a formalisation of often practiced manual intervention in computational forecasts.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022} was launched in March 2021 in order to elicit weekly predictions for various COVID-19 related forecast targets from different research groups. The forecasts evaluated in this study were submitted every Monday before 11.59pm GMT between May 24 2021 and August 16 2021. Forecasts were made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epidemiological weeks, ending on a Saturday, and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub followed a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecasts submitted to the hub were automatically checked for conformity with the required format and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble ("EuroCOVIDhub-ensemble") shown on all official Forecast Hub visualisations (https://covid19forecasthub.eu/) was a mean ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July onwards, the default Forecast Hub ensemble became a median ensemble. The median number of models included in the Forecast Hub ensemble for the UK during the study period was 9 for cases and 10 for deaths (see Figure \ref{fig:num-forecasters} in the SI).

@@

Just as in the previous study, the default forecast shown was a repetition of the last known observation with constant uncertainty around it. The shown distribution was the exponential of a normal distribution with mean $\log (\text{last value})$ and uncertainty set to the standard deviation of the last four changes in weekly log observed forecasts (i.e., as $\sigma(\log(\text{value4}) - \log(\text{value3}), \, \log(\text{value3}) - \log(\text{value2}), \, \dots)$).
In addition to information about past observations, participants could see various metrics and data such as the test positivity rate and vaccination rate sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

In addition to the 'direct' forecasts, we implemented a second forecasting method ('$R_t$ forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$. This forecast was made based on a baseline estimate produced by the \texttt{EpiNow2} \citep{epinow2} \textsf{R} \cite{R} package effective reproduction number model which we also used in \cite{bosseComparingHumanModelbased2022} as a standalone computational model. The estimate produced by \texttt{EpiNow2} was shown as the default forecast and could be adjusted by the user. The resulting $R_t$ forecast was then translated into a forecast of cases using the simulation model from the \texttt{EpiNow2} \textsf{R} package, which implements a renewal equation based \citep{fraserEstimatingIndividualHousehold2007} generative process for latent infections. We chose a Gaussian Process prior with mean 0 for the first differences of the effective reproduction number in time, implying that in the absence of informative data the reproduction number would remain constant on average, with uncertainty increasing with the temporal distance to informative data points. Latent infections were convolved with delay distributions representing the incubation period and reporting delay, and assumed to follow a negative binomial observation model with a day of the week effect to produce an estimate of reported cases. This approach has been widely used for short-term forecasting \cite{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022} and used to produce reproduction number estimates \cite{abbottEstimatingIncreaseReproduction2021, abbottEstimatingTimevaryingReproduction2020a, sherrattExploringSurveillanceData2021}. Further details are given in the SI.

To obtain forecasts for deaths, we similarly fit a model that convolved observed and predicted reported cases as implied by the $R_t$ forecast over a  delay distributions \cite{sherrattExploringSurveillanceData2021, abbottEstimatingTimevaryingReproduction2020a} and scaled them by a fixed ratio to model the time between a case report and a reported death and the case fatality ratio using the \texttt{EpiNow2} \textsf{R} package \citep{epinow2}. Further details are given in the SI and in our previous work \citep{bosseComparingHumanModelbased2022}.

@@

To boost participation compared to our last crowd forecasting study in Germany and Poland \citep{bosseComparingHumanModelbased2022} which struggled in this regard, we announced an official tournament, the "UK Crowd Forecasting Challenge". Participants were asked to submit weekly predictions for reported cases and deaths linked to COVID-19 in the United Kingdom one to four weeks into the future. Everyone who had submitted a forecast for targets in the UK during the tournament period from the 24th of May 2021 to the 16th of August 2021 was deemed a participant and eligible for a prize. The first prize was 100 GBP, second prize 50 GBP and third prize 25 GBP. Participant performance was determined using the mean weighted interval score (WIS) on the log scale (see details in the next Section), %\ref{sec:analysis}), 
averaged across forecast dates, horizons and forecast targets. For the tournament ranking, participants who did not submit a forecast in a given week were assigned the median score of all other participants who submitted a forecast that week. The UK crowd forecasting challenge was announced over Twitter and our networks. 
% (e.g. https://www.voice-global.org/public/opportunities/archived/uk-covid-19-crowd-forecasting-challenge/)
In addition, we created a project website, \url{https://crowdforecastr.org}, made weekly posts on Twitter and sent participants who had registered on the online application weekly emails with a reminder and a summary of their past performance. A public leaderboard was available on our website \url{https://epiforecasts.io}. Participants could choose to make a direct forecast as well as an $R_t$ forecast and were counted as two separate forecasters and eligible for prizes twice. Weekly forecasts had to be submitted between Sunday 12pm and Monday 8pm UK time.

@@

We scored forecasts using the weighted interval score (WIS, formulas given in the SI) 
\cite{bracherEvaluatingEpidemicForecasts2021}. The WIS is a strictly proper scoring rule yielding non-negative values, with smaller values implying better performance. A forecaster, in expectation, optimises their score by providing a predictive distribution $F$ that is equal to the data-generating distribution $G$, and is therefore incentivised to report their true belief. The WIS can be understood as an approximation of the continuous ranked probability score (CRPS) for forecasts in a quantile-based format, which itself represents a generalisation of the absolute error to predictive distributions. The WIS can be decomposed into three separate penalty components: forecast dispersion (i.e., uncertainty of forecasts), overprediction and underprediction.

\citet{bosseTransformationForecastsEvaluating2023} recently suggested to transform forecasts and observations using the natural logarithm prior to applying the WIS to better reflect the exponential nature of the underlying disease process. We, therefore, also compute WIS values after transforming all forecasts and observations using the function $f\colon x \to \log (x + 1)$. In the following, we refer to WIS scores obtained without a transformation as "scores on the natural scale", and WIS values obtained after log-transforming forecasts and observations as "scores on the log scale". To make scores easier to interpret, we report relative WIS scores, where the average score for a given model was divided by the average score for the European Forecast Hub ensemble ("EuroCOVIDhub-ensemble"). In addition, we computed ranks based on WIS values.

In order to measure probabilistic calibration \citep{gneitingProbabilisticForecastsCalibration2007}, we used the empirical coverage of all central 50\% and 90\% prediction intervals. Empirical coverage refers to the percentage of observations falling inside any given central prediction interval (e.g., the cumulative percentage of observed values that fall inside all central 50\% prediction intervals).

If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs, which found predictive performance to be poor and unreliable beyond this horizon \cite{cramerEvaluationIndividualEnsemble2021, sherrattPredictivePerformanceMultimodel2022, bracherShorttermForecastingCOVID192021}. We analysed all forecasts stratified by forecast target (cases or deaths), forecast horizon, and forecast approach. We compared the performance of the direct vs. $R_t$ forecasting approach using instances where we had both a direct forecasts and an $R_t$ forecast from the same person.

@@

A total number of 90 participants submitted forecasts (more precisely, forecasts were submitted from 90 different accounts, some of them anonymous). Out of 90 participants, 21 self-identified as "experts", i.e., stated they had professional experience in infectious disease modelling or a related field.

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/scores-and-forecasts.png}
\caption{\bf{Forecasts and corresponding WIS for 2-week ahead forecasts of cases and deaths from COVID-19 in the UK.} A: 50\% prediction intervals (coloured bars) and observed values (black line and points) for cases and deaths on the natural scale. B: Corresponding WIS values, decomposed into dispersion, overprediction and underprediction. C: 50\% prediction intervals on the log scale, i.e., after applying the natural logarithm to all forecasts and observations. D: Corresponding WIS on the log scale, i.e., the WIS applied to the log-transformed forecasts and observations.} 
\label{fig:forecasts-scores} 
\end{figure*}

@@

\subsection*{Experts and non-experts}

@@

%Calibration
In line with results from previous work \citep{bosseComparingHumanModelbased2022, sherrattPredictivePerformanceMultimodel2022}, we found almost all forecasts for cases to be underdispersed (i.e., too narrow/overconfident). Empirical coverage for death forecasts was higher than the corresponding coverage for cases for all forecasting approaches except the $R_t$ forecasts.

@@

This study explored a novel method of forecasting infectious diseases that combines a human forecast of the estimated effective reproduction number $R_t$ with epidemiological modelling to map the $R_t$ forecast to a forecast of cases and deaths. One appeal of this approach is that the forecaster can directly forecast the generative process and how they believe it is affected by interventions and changes in behaviour. Computational modelling then takes care of dealing with details such as reporting delays, generation intervals, day of the week periodicity, and the relationship between different indicators. This could help reduce cognitive load, and make it easier to synthesise various sources in information into a single forecast, at least for forecasters who have an intuitive understanding of $R_t$. Though we note all of these modelling steps and the construction of the model itself requires the human constructing the model to make assumptions. Anecdotally, forecasters familiar to the authors reported high satisfaction with the forecasting experience. In our study, $R_t$ forecasts of cases were comparable to direct forecasts, with a tendency for good forecasts to improve when using the $R_t$ method and worse forecasts to deteriorate even more. Sample sizes, however, were very low. 
Given that forecasters could simulate cases in the app, it is also possible that forecasters were in fact directly forecasting cases. 
$R_t$ forecasts of deaths (which forecasters could not see in the app) were noticeably worse than direct forecasts of deaths. The computational model underlying our $R_t$ forecasts of deaths estimated a constant CFR and delay distribution using the last 4 weeks of data, therefore updating relatively slowly to new circumstances and was assumed to be constant over the four week forecast horizon. However, as mentioned before, the CFR evolved during the study period. Forecasters had no way of inspecting the death forecast implied by their $R_t$ forecast and also had no way to adjust the CFR manually, likely impacting forecast accuracy. Giving human forecasters the ability to adjust the CFR and other model parameters would have increased complexity of the interface, but would have solved issues with the assumptions of the underlying model. Alternatively, a more complex model could have been used which allowed for time-varying CFR estimates and forecast these changes over the forecast horizon though this approach may still have struggled to cope with the rapid changes observed during the study period. Another important limitation is that we didn't have full sample trajectories of the $R_t$-values predicted by forecasters. Rather, trajectories had to be constructed based on the distributions provided for the different forecast horizons, which likely negatively affected forecasts. One potential way to disentangle the effect of the convolution model from the $R_t$ forecasts would have been to use the human forecasts for cases as an input to the second computational model, which could then have simulated deaths. Future work could expose forecasters to different combinations of these options with the aim of separating effects of the user interface from ones related to the structure of the underlying computational model.

@@

Past expert forecasts of COVID-19 \cite{recchiaHowWellDid2021} had found predictions from experts to outperform those of non-experts. In our study, an ensemble of self-reported "experts" outperformed an ensemble of "non-experts" when forecasting cases two weeks ahead, both on the log scale and on the natural scale. When forecasting deaths two weeks ahead, "experts" performed worse than "non-experts" on the log scale, but better on the natural scale. Forecasts for "experts" tended to be better calibrated than non-experts.  However results should be taken with care considering relatively low sample sizes (median of 11 "non-experts" for cases and 10 for deaths, median of 8 "experts" for cases and deaths) and given that expert status was self-reported. Furthermore, we only asked for professional involvement in a field related to infectious disease modelling, not specifically for familiarity with modelling of COVID-19 in the UK, and only offered participants a binary choice. However, as we used ad-hoc recruitment in our networks many of these self-identified experts are likely to be infectious disease modellers.

% \paragraph{Susceptibility of results}
Overall, results of our study should be taken with caution due to several important limitations. Firstly, our study was restricted to one location and to a relatively short period of thirteen weeks. Secondly, there were many confounding factors that likely influence results. These include the fact that different participants made forecasts at different points in time (with the median forecaster only submitting a single forecast) and that subgroups of interest (e.g. "experts", or $R_t$ forecasts) had different numbers of forecasters. 
In most instances, differences in scores between forecast approaches were small compared to the variance of scores within a single approach. In addition, there were many researcher degrees of freedom that could influence findings, for example how individual forecasts were combined to create an ensemble. Results were influenced by choices made during the evaluation with, for example, some conclusions  depending on forecast horizon and the transformation used prior to scoring. 
Highlighting this, prizes to the human forecasters were paid out based on the combined WIS on the log scale across all horizons and forecast targets. Had we chosen to instead measure WIS on the natural scale, or to forecast only cases and continue to score on the log scale, rankings and pay-outs would have been different.

@@

\subsection*{Author contributions}
NIB contributed to the conceptualization, data curation, formal analysis, investigation, methodology, software development, validation, visualization, and original draft preparation of the manuscript, as well as its review and editing.
SA contributed to the conceptualization, data curation, investigation, methodology, software development, supervision, and review and editing of the manuscript.
JB contributed to the supervision, review, and editing of the manuscript.
EvL contributed to the conceptualization, supervision, and review and editing of the manuscript.
AC contributed to the conceptualization, supervision, and review and editing of the manuscript.
SF contributed to the conceptualization, funding acquisition, project administration, supervision, and review and editing of the manuscript.
All authors have read and approved the final version of the manuscript.

\subsection*{Competing interests}
NIB is a part-time employee of Metaculus, an online prediction platform.
